---
title: "R Notebook"
output: html_notebook
---


# General stuff


```{r}
install.packages("tidyverse")
library(dplyr)
library(ggplot2)

install.packages("caret")
library(caret)

```

```{r}


std_dev_position <- function(X, Y) {
  x_mean <- mean(X)
  y_mean <- mean(Y)
  n <- length(X)
  
  return(sqrt(1/(n-1) * sum((X - x_mean)^2 + (Y - y_mean)^2)))
}


```



# Experiment 1: Personal-confidence effect on clustering

## Shubert
```{r}

## Shubert
shubert <- read.csv("netlogo_personal_confidence_penalty_constraints_shubert_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "personal_confidence", "x", "y", "personal_best_val", "global_best_val"))

dispersion_shubert_pos <- shubert %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_position = std_dev_position(x,y))

dispersion_shubert_best_val <- shubert %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_best_val = sd(personal_best_val)) 

lm_shubert <- lm(std_dev_best_val ~ personal_confidence, dispersion_shubert_best_val)
summary(lm_shubert)


ggplot(dispersion_shubert_pos, aes(x=personal_confidence, y=std_dev_position)) + 
  geom_point(aes(colour = fitness_function))


```

## Schaffer
```{r}
schaffer <- read.csv("netlogo_personal_confidence_penalty_constraints_schaffer_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "personal_confidence", "x", "y", "personal_best_val", "global_best_val"))

dispersion_schaffer_pos <- schaffer %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_position = std_dev_position(x,y)) 

dispersion_schaffer_best_val <- schaffer %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_best_val = sd(personal_best_val))

lm_schaffer <- lm(std_dev_position ~ personal_confidence, dispersion_schaffer_pos)
summary(lm_schaffer)


ggplot(dispersion_schaffer_best_val, aes(x=personal_confidence, y=std_dev_best_val)) + 
  geom_point(aes(colour = fitness_function))

```




## Eggholder
```{r}
eggholder <- read.csv("netlogo_personal_confidence_penalty_constraints_eggholder_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "personal_confidence", "x", "y", "personal_best_val", "global_best_val"))

dispersion_eggholder_pos <- eggholder %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_position = std_dev_position(x,y))

dispersion_eggholder_best_val <- eggholder %>% group_by(fitness_function, personal_confidence, run_nr) %>% summarize(std_dev_best_val = sd(personal_best_val))

lm_eggholder <- lm(std_dev_best_val ~ personal_confidence, dispersion_eggholder_best_val)
summary(lm_eggholder)


ggplot(dispersion_eggholder_pos, aes(x=personal_confidence, y=std_dev_position)) + 
  geom_point(aes(colour = fitness_function))

```





## Plotting together standard deviation position
```{r}
dispersion_all_pos <- bind_rows(dispersion_schaffer_pos, dispersion_shubert_pos, dispersion_eggholder_pos)
dispersion_all_pos$fitness_function <- factor(dispersion_all_pos$fitness_function, 
                                          levels=c("Fitness function Schaffer", "Fitness function Shubert", "Fitness function Eggholder"),
                                          labels=c("Schaffer", "Shubert", "Eggholder"))

lm_all <- lm(std_dev_position ~ personal_confidence + personal_confidence:fitness_function, dispersion_all_pos)
summary(lm_all)

ggplot(dispersion_all_pos, aes(x=personal_confidence, y=std_dev_position)) + 
  geom_point(aes(colour = fitness_function))

```



## Plotting together standard deviation personal best value
```{r}
dispersion_all_best_val <- bind_rows(dispersion_schaffer_best_val, dispersion_shubert_best_val, dispersion_eggholder_best_val)
dispersion_all_best_val$fitness_function <- factor(dispersion_all_best_val$fitness_function, 
                                          levels=c("Fitness function Schaffer", "Fitness function Shubert", "Fitness function Eggholder"),
                                          labels=c("Schaffer", "Shubert", "Eggholder"))

lm_all <- lm(std_dev_best_val ~ personal_confidence , dispersion_all_best_val)
summary(lm_all)

ggplot(dispersion_all_best_val, aes(x=personal_confidence, y=std_dev_best_val)) + 
  geom_point(aes(colour = fitness_function))

```

standard deviation personal best value with interaction between fitness functions
```{r}
lm_all <- lm(std_dev_best_val ~ personal_confidence + personal_confidence:fitness_function, dispersion_all_best_val)
summary(lm_all)

```





# Experiment 2: Convergence after iterations number

## Schaffer stop
```{r}
schaffer <- read.csv("netlogo_speed_limit_penalty_constraints_schaffer_stop_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "speed_limit", "x", "y", "personal_best_val", "iterations"))

dispersion_schaffer <- schaffer %>% group_by(fitness_function, speed_limit, run_nr) %>% summarize(avg_iter = mean(iterations)) 

lm_schaffer <- lm(avg_iter ~ speed_limit, dispersion_schaffer)
summary(lm_schaffer)


ggplot(dispersion_schaffer, aes(x=speed_limit, y=avg_iter)) + 
  geom_point(aes(colour = fitness_function))

```



## eggholder stop
```{r}
eggholder <- read.csv("netlogo_speed_limit_penalty_constraints_eggholder_stop_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "speed_limit", "x", "y", "personal_best_val", "iterations"))

dispersion_eggholder <- eggholder %>% group_by(fitness_function, speed_limit) %>% summarize(avg_iter = mean(iterations)) 

lm_eggholder <- lm(avg_iter ~ speed_limit, dispersion_eggholder)
summary(lm_eggholder)


ggplot(dispersion_eggholder, aes(x=speed_limit, y=avg_iter)) + 
  geom_point(aes(colour = fitness_function))

```

## Shubert stop
```{r}
shubert <- read.csv("netlogo_speed_limit_penalty_constraints_shubert_stop_coordinates.csv", 
                    header = FALSE, col.names = c("fitness_function", "run_nr", "speed_limit", "x", "y", "personal_best_val", "iterations"))

dispersion_shubert <- shubert %>% group_by(fitness_function, speed_limit, run_nr) %>% summarize(avg_iter = mean(iterations)) 

lm_shubert <- lm(avg_iter ~ speed_limit, dispersion_shubert)
summary(lm_shubert)


ggplot(dispersion_shubert, aes(x=speed_limit, y=avg_iter)) + 
  geom_point(aes(colour = fitness_function))

```




```{r}
dispersion_all <- bind_rows(dispersion_schaffer, dispersion_shubert, dispersion_eggholder)
dispersion_all$fitness_function <- factor(dispersion_all$fitness_function, 
                                          levels=c("Fitness function Schaffer", "Fitness function Shubert", "Fitness function Eggholder"),
                                          labels=c("Schaffer", "Shubert", "Eggholder"))

lm_all <- lm(avg_iter ~ speed_limit + speed_limit:fitness_function, dispersion_all)
summary(lm_all)

ggplot(dispersion_all, aes(x=speed_limit, y=avg_iter)) + 
  geom_point(aes(colour = fitness_function))


```




experiment2 <- read.csv("netlogo_population_size_best_vals.csv", 
                        header = FALSE, col.names = c("run_nr", "fitness_function", "population_size", "global_best_val"))

experiment2$fitness_function <- factor(experiment2$fitness_function, 
                                       levels=c("Fitness function Schaffer", "Fitness function Shubert", "Fitness function Eggholder"),
                                       labels=c("Schaffer", "Shubert", "Eggholder"))

lm_experiment2 <- lm(global_best_val ~ population_size, experiment2)
summary(lm_experiment2)

ggplot(experiment2, aes(x=population_size, y=global_best_val)) + 
  geom_point(aes(colour = fitness_function), position = position_jitter(width = 1, height = 0.02))


